Staged Abstract Interpreters

### 2019-09-05 paper/70_eval.tex
Proofreading eval

To evaluate how performance can be improved by our staged abstract interpreters
approach, we scale our toy interpreter to a large subset of the Scheme language
and evaluate the analyzing time against with the unstaged versions. The result
shows that the staged versions are averagely 10+ times as fast as the unstaged
version, as evidence of our \textit{abstraction without regret} approach.

\paragraph{Implementation and Evaluation Environment} We implement the abstract
interpreters for control-flow analysis, i.e., a value-flow analysis that
computes which lambda terms can be called at each call-site. 
With other suitable abstract domains, the abstract interpreter
can also be extended to other analyses. 
A front-end desugars Scheme to a small core language that makes the analyzer
easy to implement. We implement two common monovariant 0-CFA-like analysis, one
is equipped with store-widening (0CFA-SW for short), the other does not (0CFA
for short). Usually, the store-widened analysis is expected to be faster than the one
without.  Our prototype implementation currently generates Scala code; the
generated Scala code will be compiled and executed on JVM. It is possible to
generate C/C++ code in the future. In the experiments, we implement several
optimizations mentioned in Section \ref{staged_ds}, specifically, the selective
caching scheme and IR-level rewriting rules (for pairs, lists, sets, and maps)
that exploit partially-static data. These optimizations are useful to reduce
the size of generated code.

We use Scala 2.12.8 and Oracle JVM 1.8.0-191 \footnote{The options for
running JVM is set to the following: \texttt{-Xms2G -Xmx8G -Xss1024M
-XX:MaxMetaspaceSize=2G - XX:ReservedCodeCacheSize=2048M}},
running on an Ubuntu 16.04 LTS (kernel 4.4.0) machine. All of our
evaluations were performed on a machine with 4 Intel Xeon Platinum
8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96 cores
(192 threads) in total, the abstract interpreters are single-threaded.
To minimize the effect of warm-up from the HotSpot JIT compiler,
all the experiments are executed for 20 times and we report the
statistical median value of the running times. We set a 5 minutes
timeout.

\paragraph{Benchmarks}
The benchmark programs we used in the experiment are collected from
previous papers \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
DBLP:journals/corr/abs-1102-3676} and existing artifacts
\footnote{https://github.com/ilyasergey/reachability} for control-flow
analysis.  Some of the benchmarks are small programs designed for
experiments, such as the @kcfa-worst-@$n$ series, which are intended
to challenge $k$-CFA; while some other benchmarks are from real-world
applications, for examples, the RSA public key encryption algorithm
@rsa@. In Figure \ref{evaluation_result}, we report the number of
AST nodes after parsing and desugaring (excluding comments) as a proper
measurement of program size.

@@

Figure~\ref{evaluation_result} shows the evaluation result, comparing the
performance improved by staging on the two monovariant CFAs. The
\textit{unstaged} and \textit{staged} columns show the median time (excluding
pre-compilation) to finish the analysis. The column
$\frac{\text{unstaged}}{\text{staged}}$ shows the improvement. The dash @-@ in
some cells indicates timeout.  As expected, the staged version significantly
outperforms the unstaged version on all benchmarks. 
%We observe that simple value domains used in the benchmark usually leads to
%better speed-ups, for example, the @kcfa-worst@ series only uses lambdas and
%booleans.
For 0CFA w/o store-widening, we observe an average speed-up of ~17x times; for
0CFA w/ store-widening, the average speed-up is ~7x times.
\texttt{kcfa-worst-64} has the most speed-up among all the tests.
Although it has a comparably large size of AST, it has a quite simple structure
which only involves booleans, lambda terms and applications. We conjecture that
the simple domain and structure allow the generated code to be efficiently
optimized by the JVM. But on the corresponding 0CFA-SW, we do not observe such
great speed-up. Due to the complexity of generated code and next-stage running
platform (JVM), we leave the detailed empirical analysis of which static
analysis can benefit more from the specialization as future work.

>>>>

To evaluate how performance can be improved by our staged abstract interpreters
approach, we scale our toy interpreter to a large subset of the Scheme language
and evaluate the analysis time against the unstaged versions. The result
shows that the staged versions are, on average, 10+ times as fast as the unstaged
version, as evidence of our \textit{abstraction without regret} approach.

\paragraph{Implementation and Evaluation Environment} We implement the abstract
interpreters for control-flow analysis, i.e., a value-flow analysis that
computes which lambda terms can be called at each call-site.
With other suitable abstract domains, the abstract interpreter
can also be extended to other analyses.
A front-end desugars Scheme to a small core language that makes the analyzer
easy to implement. We implement two common monovariant 0-CFA-like analyses: one
is equipped with store-widening (0CFA-SW for short), the other is not (0CFA
for short). It is traditionally expected that the store-widened analysis be faster than the
alternative.  Our prototype implementation currently generates Scala code; the
generated Scala code will be compiled and executed on the JVM, though it is possible to
generate C/C++ code in the future. In the experiments, we implement several
optimizations mentioned in Section \ref{staged_ds}, specifically, the selective
caching scheme and IR-level rewriting rules (for pairs, lists, sets, and maps)
that exploit partially-static data. These optimizations are useful to reduce
the size of generated code.

We use Scala 2.12.8 and Oracle JVM 1.8.0-191 \footnote{The options for
running the JVM are set to the following: \texttt{-Xms2G -Xmx8G -Xss1024M
-XX:MaxMetaspaceSize=2G - XX:ReservedCodeCacheSize=2048M}},
running on an Ubuntu 16.04 LTS (kernel 4.4.0) machine. All of our
evaluations were performed on a machine with 4 Intel Xeon Platinum
8168 CPU at 2.7GHz and 3 TB of RAM. Although the machine has 96 cores
(192 threads) in total, the abstract interpreters are single-threaded.
To minimize the effect of warm-up from the HotSpot JIT compiler,
all the experiments are executed for 20 times and we report the
statistical median value of the running times. We set a 5 minute
timeout.

\paragraph{Benchmarks}
The benchmark programs we used in the experiment are collected from
previous papers \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
DBLP:journals/corr/abs-1102-3676} and existing artifacts
\footnote{https://github.com/ilyasergey/reachability} for control-flow
analysis.  Some of the benchmarks are small programs designed for
experiments, such as the @kcfa-worst-@$n$ series, which are intended
to challenge $k$-CFA; while some other benchmarks are from real-world
applications, for example, the RSA public key encryption algorithm
@rsa@. In Figure \ref{evaluation_result}, we report the number of
AST nodes after parsing and desugaring (excluding comments) as a proper
measurement of program size.

@@

Figure~\ref{evaluation_result} shows the evaluation result, comparing the
performance improved by staging on the two monovariant CFAs. The
\textit{unstaged} and \textit{staged} columns show the median time (excluding
pre-compilation) to finish the analysis. The column
$\frac{\text{unstaged}}{\text{staged}}$ shows the improvement. The dash @-@ in
some cells indicates timeout.  As expected, the staged version significantly
outperforms the unstaged version on all benchmarks.
%We observe that simple value domains used in the benchmark usually leads to
%better speed-ups, for example, the @kcfa-worst@ series only uses lambdas and
%booleans.
For 0CFA without store-widening, we observe an average speedup of ~17x times;
for 0CFA with store-widening, the average speedup is ~7x times.
\texttt{kcfa-worst-64} has the highest speedup among all the tests. Although it
has a comparably large size of AST, it has a rather simple structure which only
involves booleans, lambda terms, and applications. We conjecture that the simple
domain and structure allow the generated code to be efficiently optimized by the
JVM. However, on the corresponding 0CFA-SW, we do not observe such a significant
speedup. Due to the complexity of generated code and next-stage running platform
(JVM), we leave the detailed empirical analysis of which static analysis can
benefit more from the specialization as future work.
